\textbf{SPECFEM3D}\\ \textbf{User Manual}

### 2010-11-04 doc/USER_MANUAL/manual_SPECFEM3D.tex
added the source code of SCOTCH; updated the manual (history, citations, SCOTCH).

The software package SPECFEM3D simulates seismic
wave propagation at the local or regional scale
based upon the spectral-element method (SEM).
The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
Note that in many (most?) geological models in the context of seismic wave propagation studies
a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient.
Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. For a detailed introduction to the SEM as applied to regional
seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
in particular \citet{KoLiTrSuStSh04}. If you use the 3D southern
California model, please cite \citet{SuSh03} (LA), \citet{lovelyetal06}
(Salton Trough), and \citet{hauksson2000} (southern California).
The Moho map was determined by \citet{zhu&kanamori2000}. The 1D SoCal
model was developed by \citet{DrHe90}. The package can accommodate
full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
variations in attenuation. Adjoint capabilities and finite-frequency
kernel simulations are included \citep{LiTr06,TrKoLi08}.

@@

SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
conference in Phoenix, Arizona (USA) by running at 5 teraflops (sustained)
on 1944 processors of the Japanese Earth Simulator using 14.6 billion
degrees of freedom stored in 2.5 terabytes of memory; see \cite{KoTsChTr03}
and the Gordon Bell Awards News Release \url{www.sc-conference.org/sc2003/nr_finalaward.html }
for details. It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.

@@

@@

This material is based upon work supported by the USA National Science
Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
Reintegration Grant No. MIRG-CT-2005-017461. 
Older versions of the code were initially developed by Dimitri Komatitsch at Institut de Physique du Globe (France)
and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University (USA).
Any opinions, findings,
and conclusions or recommendations expressed in this material are
those of the authors and do not necessarily reflect the views of the
USA National Science Foundation, CNRS, INRIA, ANR or the European
Marie Curie program.

@@

An efficient partitioning is important because it leverages the overall running time of the application.
It amounts to balance the number of elements in each slice while minimizing the communication coasts
resulting from the placement of adjacent elements to different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the Scotch library
which provides efficent static mapping, graph and mesh partitioning routines. Scotch is a free software package downloadable from the web page \texttt{https://gforge.inria.fr/projects/scotch/}.
\\

Prior to compiling \texttt{decompose\_mesh\_SCOTCH}, make sure you have correctly specified the path 
of the Scotch library with the option \texttt{-{}-with-scotch-lib} of the \texttt{configure} script. Then you are ready to compile, 
in the directory \texttt{decompose\_mesh\_SCOTCH} type '\texttt{make decompose\_mesh\_SCOTCH}'.
If all paths and flags have been set correctly, the executable \texttt{xdecompose\_mesh\_SCOTCH} should be produced.

The partitioning is done in serial, 
the synopsis is :
\begin{lyxcode}
./decompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
\end{lyxcode}

>>>>

The software package SPECFEM3D simulates seismic
wave propagation at the local or regional scale
based upon the spectral-element method (SEM).
The SEM is a continuous Galerkin technique, which can easily be made discontinous \citep{BeMaPa94,KoWoHu02,ChCaVi03,Kop06}; it is then a particular case of the discontinuous Galerkin technique \citep{ReHi73,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,BeLaPi06,DuKa06,DeSeWh08,WiStBuGh10,DeSe10,EtChViGl10}, with optimized efficiency because of its tensorized basis functions.
Note that in many (most?) geological models in the context of seismic wave propagation studies
(except for fault dynamic rupture studies)
a discontinous mesh is not needed because material property contrasts are not drastic and thus a continuous formulation is sufficient.
Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. For a detailed introduction to the SEM as applied to regional
seismic wave propagation, please consult \citet{KoVi98,KoTr99,ChKoViCaVaFe07,TrKoLi08} and
in particular \citet{KoLiTrSuStSh04}. The package can accommodate
full 21-parameter anisotropy (see~\citet{ChTr07}) as well as lateral
variations in attenuation. Adjoint capabilities and finite-frequency
kernel simulations are included \citep{LiTr06,TrKoLi08}.

The SPECFEM3D package was first developed by Dimitri Komatitsch and Jeroen Tromp at Harvard University, then Caltech and Princeton, USA, starting in 1998, based in part on earlier work by Dimitri Komatitsch and Jean-Pierre Vilotte at IPG in Paris, France from 1994 to 1997. Since then it has been developed and maintained by a development team: in alphabetical order, C\'eline Blitz, Emiljana Jorgji, Dimitri Komatitsch, Nicolas Le Goff, Pieyre Le Loher, Roland Martin, David Mich\'ea, Jeroen Tromp \toall{add full list of developers here}.

@@

SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
conference in Phoenix, Arizona (USA) by running at 5 teraflops (sustained)
on 1944 processors of the Japanese Earth Simulator using 14.6 billion
degrees of freedom stored in 2.5 terabytes of memory; see \cite{KoTsChTr03}
and the Gordon Bell Awards News Release \url{www.sc-conference.org/sc2003/nr_finalaward.html }
for details. It was a finalist again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories (USA) \citep{CaKoLaTiMiLeSnTr08}.
It also won the BULL Joseph Fourier supercomputing award in 2010.

@@

If you use the 3D southern
California model, please cite \citet{SuSh03} (Los Angeles model), \citet{lovelyetal06}
(Salton Trough), and \citet{hauksson2000} (southern California).
The Moho map was determined by \citet{zhu&kanamori2000}. The 1D SoCal
model was developed by \citet{DrHe90}.

@@

This material is based upon work supported by the USA National Science
Foundation under Grants No. EAR-0406751 and EAR-0711177, by the French
CNRS, French INRIA Sud-Ouest MAGIQUE-3D, French ANR NUMASIS under
Grant No. ANR-05-CIGC-002, and European FP6 Marie Curie International
Reintegration Grant No. MIRG-CT-2005-017461.
Any opinions, findings,
and conclusions or recommendations expressed in this material are
those of the authors and do not necessarily reflect the views of the
USA National Science Foundation, CNRS, INRIA, ANR or the European
Marie Curie program.

@@

An efficient partitioning is important because it leverages the overall running time of the application.
It amounts to balance the number of elements in each slice while minimizing the communication coasts
resulting from the placement of adjacent elements to different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the Scotch library
which provides efficent static mapping, graph and mesh partitioning routines. Scotch is a free software package developed by
Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \texttt{https://gforge.inria.fr/projects/scotch/}. It is more recent than METIS, actively maintained and performs better in many cases.
\\

Prior to compiling \texttt{decompose\_mesh\_SCOTCH}, make sure you have correctly specified the path
of the Scotch library with the option \texttt{-{}-with-scotch-lib} of the \texttt{configure} script.
For convenience we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
in directory src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b. Please refer to file INSTALL.txt in that directory to see how to compile it.
In the future you should be able to find for recent versions at http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html .
\\

Then you are ready to compile,
in the directory \texttt{decompose\_mesh\_SCOTCH} type '\texttt{make decompose\_mesh\_SCOTCH}'.
If all paths and flags have been set correctly, the executable \texttt{xdecompose\_mesh\_SCOTCH} should be produced.

The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
the synopsis is:
\begin{lyxcode}
./decompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
\end{lyxcode}
